---
title: "Georgia Run-Off 2022"
subtitle: "How Do Political Campaigns Target Voters?"
author: "Who Targets Me<br>Fabio Votta"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
    highlight: kate
---



```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(
    cache = T, 
    echo = F, 
    warning = F, 
    message = F, 
    cache.lazy = FALSE
)


# pacman::p_load(tidyverse, highcharter)
library(tidyverse)
library(highcharter)
library(gt)



options(scipen = 999)

source("helpers.R")

# pro_reps <- us_advertisers %>% 
#   filter(left_vs_right == "All Pro-Republican Pages")

# source("utils.R")
```

```{r}

us_advertisers <- vroom::vroom("data/wtm-advertisers-us-2022-11-28T14_22_01.338Z.csv") %>%
  # count(entities.short_name, sort = T) %>%
  filter(entities.short_name %in% c("GOP", "Dems", "DemPACs", "RepPACs", "Con", "Prog")) %>%
  mutate(left_vs_right = ifelse(entities.short_name %in% c("GOP", "Con", "RepPACs"), "All Pro-Republican Pages", "All Pro-Democrat Pages")) %>%
  select(page_name = name, page_id = advertisers_platforms.advertiser_platform_ref, advertiser_type = entities.short_name, left_vs_right) %>%
  mutate(advertiser_type = case_when(
    advertiser_type == "Con" ~ "Conservatives (non-party, non-PACs)",
    advertiser_type == "Prog" ~ "Progressives  (non-party, non-PACs)",
    advertiser_type == "RepPACs" ~ "Republican PACs",
    advertiser_type == "DemPACs" ~ "Democrat PACs",
    advertiser_type == "GOP" ~ "GOP (candidates, parties)",
    advertiser_type == "Dems" ~ "Democrats (candidates, parties)"
  )) #%>% 

runoff <- readRDS("data/runoff7.rds")  %>%
  left_join(us_advertisers %>% mutate(internal_id = as.character(page_id))) %>%
  mutate(total_spend_formatted = readr::parse_number(total_spend_formatted)) %>%
  mutate(total_spend_formatted = ifelse(total_spend_formatted == 100, 1, total_spend_formatted)) %>%
  mutate(total_spend = total_spend_formatted * total_spend_pct) 

runoff30 <- readRDS("data/runoff.rds")  %>%
  left_join(us_advertisers %>% mutate(internal_id = as.character(page_id))) %>%
  mutate(total_spend_formatted = readr::parse_number(total_spend_formatted)) %>%
  mutate(total_spend_formatted = ifelse(total_spend_formatted == 100, 1, total_spend_formatted)) %>%
  mutate(total_spend = total_spend_formatted * total_spend_pct) 

```


## Methodology

We annotated over 3000 US political advertisers during the 2022 Midterms to try and better understand how campaigns are targeting their ads on Meta platforms (which include Facebook and Instagram). The data source is the [Meta Ad Library](https://www.facebook.com/ads/library/), most of it taken from the new Audience Tab which shows how much money parties spend on audiences.

**Here we zoom in on the Georgia Senate runoff race** by only keeping advertisers who:

1. advertised in the last 7 days (Nov 25th - 1st December 2022)
2. advertised in the last 30 days (Nov 2nd - 1st December 2022) but only if they spend any money on the state of Georgia

> Note: Meta only provides 7, 30 and 90 days window for this data and their data lags behind. We will update the report as soon as new data is available. 






## Overall Stats  {.tabset .tabset-fade .tabset-pills}

First, some overall statistics. Pro-Democrat pages are clearly and consistently outspending pro-Republican pages. Also in terms of volume, Pro-Democrat pages run considerably more ads.


### Nov 25th - 1st December 2022 (Last 7 Days)

```{r}
runoff %>% 
  distinct(internal_id, .keep_all = T) %>% 
  group_by(left_vs_right) %>% 
  summarize(total_num_ads = n()) %>% 
  drop_na() %>% 
    mutate(total_num_ads = scales::comma(total_num_ads)) %>%
  pivot_wider(names_from = left_vs_right, values_from = total_num_ads) %>% 
  mutate(type = "Total Number of Advertisers") %>% 
  
  
  bind_rows(
    runoff %>% 
  distinct(internal_id, .keep_all = T) %>% 
  group_by(left_vs_right) %>% 
  arrange(desc(total_spend_formatted)) %>% 
    slice(1:3) %>% 
    mutate(total_spend_formatted = scales::comma(total_spend_formatted)) %>%
    mutate(lab = paste0(page_name, " ($", total_spend_formatted, ")")) %>% 
    select(left_vs_right, lab) %>% 
  drop_na() %>% 
  summarize(lab = paste0("<br>", 1:3, ". ", lab, collapse = "")) %>% 
  pivot_wider(names_from = left_vs_right, values_from = lab) %>% 
  mutate(type = "Top Advertiser")
  ) %>% 
  
  bind_rows(
    runoff %>% 
  distinct(internal_id, .keep_all = T) %>% 
  group_by(left_vs_right) %>% 
  summarize(total_num_ads = sum(total_num_ads)) %>% 
  drop_na() %>% 
    mutate(total_num_ads = scales::comma(total_num_ads)) %>% 
  pivot_wider(names_from = left_vs_right, values_from = total_num_ads) %>% 
  mutate(type = "Total Number of Ads")
  )%>% 
  bind_rows(
    runoff %>% 
  distinct(internal_id, .keep_all = T) %>% 
  group_by(left_vs_right) %>% 
  summarize(total_spend_formatted = sum(total_spend_formatted)) %>% 
    mutate(total_spend_formatted = scales::comma(total_spend_formatted)) %>% 
  drop_na() %>% 
  pivot_wider(names_from = left_vs_right, values_from = total_spend_formatted) %>% 
  mutate(type = "Total Spend ($)") 
  ) %>% 



  gt(
    rowname_col = "type"
    # groupname_col = "group"
  ) %>% 
  fmt_markdown(columns = everything())
```

### Nov 2nd - 1st December 2022 (Last 30 Days)

```{r}
runoff30 %>% 
  distinct(internal_id, .keep_all = T) %>% 
  group_by(left_vs_right) %>% 
  summarize(total_num_ads = n()) %>% 
  drop_na() %>% 
    mutate(total_num_ads = scales::comma(total_num_ads)) %>%
  pivot_wider(names_from = left_vs_right, values_from = total_num_ads) %>% 
  mutate(type = "Total Number of Advertisers") %>% 
  
  
  bind_rows(
    runoff30 %>% 
  distinct(internal_id, .keep_all = T) %>% 
  group_by(left_vs_right) %>% 
  arrange(desc(total_spend_formatted)) %>% 
    slice(1:3) %>% 
    mutate(total_spend_formatted = scales::comma(total_spend_formatted)) %>%
    mutate(lab = paste0(page_name, " ($", total_spend_formatted, ")")) %>% 
    select(left_vs_right, lab) %>% 
  drop_na() %>% 
  summarize(lab = paste0("<br>", 1:3, ". ", lab, collapse = "")) %>% 
  pivot_wider(names_from = left_vs_right, values_from = lab) %>% 
  mutate(type = "Top Advertiser")
  ) %>% 
  
  bind_rows(
    runoff30 %>% 
  distinct(internal_id, .keep_all = T) %>% 
  group_by(left_vs_right) %>% 
  summarize(total_num_ads = sum(total_num_ads)) %>% 
  drop_na() %>% 
    mutate(total_num_ads = scales::comma(total_num_ads)) %>% 
  pivot_wider(names_from = left_vs_right, values_from = total_num_ads) %>% 
  mutate(type = "Total Number of Ads")
  )%>% 
  bind_rows(
    runoff30 %>% 
  distinct(internal_id, .keep_all = T) %>% 
  group_by(left_vs_right) %>% 
  summarize(total_spend_formatted = sum(total_spend_formatted)) %>% 
    mutate(total_spend_formatted = scales::comma(total_spend_formatted)) %>% 
  drop_na() %>% 
  pivot_wider(names_from = left_vs_right, values_from = total_spend_formatted) %>% 
  mutate(type = "Total Spend") 
  ) %>% 



  gt(
    rowname_col = "type"
    # groupname_col = "group"
  ) %>% 
  fmt_markdown(columns = everything())
```



## Spending per Targeting Criteria  {.tabset .tabset-fade .tabset-pills}


How much Money did Campaigns Spend on Targeting Criteria? This graph shows in percent how much of their budget Pro-Democrat and Pro-Republican pages have spent on each targeting criterion. **Notably, most advertisers spend the vast majority of their budget (~65%+) on custom audiences**, which include custom E-mail lists uploaded by the advertisers.

### Nov 25th - 1st December 2022 (Last 7 Days)


```{r}

pro_reps <- us_advertisers %>% 
  filter(left_vs_right == "All Pro-Republican Pages")

pro_dems <- us_advertisers %>% 
  filter(left_vs_right == "All Pro-Democrat Pages")


runoff_spend_reps <- calculate_spend_on_each(runoff %>% filter(internal_id %in% pro_reps$page_id))

runoff_spend_dems <- calculate_spend_on_each(runoff %>% filter(internal_id %in% pro_dems$page_id))

runoff_spend_reps %>% 
  mutate(party= "Pro-Republican Pages") %>% 
  bind_rows(runoff_spend_dems %>% 
  mutate(party= "Pro-Democrat Pages")) %>% 
  filter(perc >= 0.1) %>%
  add_count(target) %>% 
  filter(n == 2) %>% 
  mutate(target = case_when(
    target == "custom_audience" ~ "Custom Audiences",
    target == "countries" ~ "GEOGRAPHY: Entire US",
    target == "regions" ~ "GEOGRAPHY: State(s)",
    target == "lookalike_audience" ~ "Lookalike Audiences",
    target == "interest" ~ "Interests",
    target == "age" ~ "Age",
    target == "zips" ~ "GEOGRAPHY: ZIP-Codes",
    target == "CITY" ~ "GEOGRAPHY: Cities",
    target == "language" ~ "Language",
    target == "gender" ~ "Gender",
    target == "electoral_districts" ~ "GEOGRAPHY: Electoral Districts",
    target == "COUNTY" ~ "GEOGRAPHY: Counties",
  )) %>% 
  mutate(target = fct_reorder(target, perc)) %>% 
  ggplot(aes(target, perc)) +
  geom_col(position = position_dodge(width = 0.9), aes(fill = party)) +
  coord_flip() +
  scale_fill_manual(name = "Pages", values = c("#008fd5", "#ff2700")) +
  theme_minimal() +
  scale_y_continuous(breaks = seq(0, 100, 10), labels = function(x) paste0(x, "%")) +
  geom_text(size = 2.5,
             aes(y = perc + 3, label = paste0(round(perc, 1), "%"), group = party),
             position = position_dodge(width = 0.9)) +
  theme(legend.position = "bottom") +
  labs(x = "", y = "\nBudget Spend on Targeting Criterias (in % of Total Budget)", caption = "Source: Meta Ad Library and data compiled by Who Targets Me. Data Viz: Fabio Votta (@favstats).")#,
       # title = "How much Money did Campaigns Spend on Targeting Criteria?",
       # subtitle = "Time Range: Nov 25th - 1st December 2022 (Last 7 Days)")

```

### Nov 2nd - 1st December 2022 (Last 30 Days)

```{r}
runoff30_spend_reps <- calculate_spend_on_each(runoff30 %>% filter(internal_id %in% pro_reps$page_id))

runoff30_spend_dems <- calculate_spend_on_each(runoff30 %>% filter(internal_id %in% pro_dems$page_id))

runoff30_spend_reps %>% 
  mutate(party= "Pro-Republican Pages") %>% 
  bind_rows(runoff30_spend_dems %>% 
  mutate(party= "Pro-Democrat Pages")) %>% 
  filter(perc >= 0.1) %>%
  add_count(target) %>% 
  filter(n == 2) %>% 
  mutate(target = case_when(
    target == "custom_audience" ~ "Custom Audiences",
    target == "countries" ~ "GEOGRAPHY: Entire US",
    target == "regions" ~ "GEOGRAPHY: State(s)",
    target == "lookalike_audience" ~ "Lookalike Audiences",
    target == "interest" ~ "Interests",
    target == "age" ~ "Age",
    target == "zips" ~ "GEOGRAPHY: ZIP-Codes",
    target == "CITY" ~ "GEOGRAPHY: Cities",
    target == "language" ~ "Language",
    target == "gender" ~ "Gender",
    target == "electoral_districts" ~ "GEOGRAPHY: Electoral Districts",
    target == "COUNTY" ~ "GEOGRAPHY: Counties",
  )) %>% 
  mutate(target = fct_reorder(target, perc)) %>% 
  drop_na() %>% 
  ggplot(aes(target, perc)) +
  geom_col(position = position_dodge(width = 0.9), aes(fill = party)) +
  coord_flip() +
  scale_fill_manual(name = "Pages", values = c("#008fd5", "#ff2700")) +
  theme_minimal() +
  scale_y_continuous(breaks = seq(0, 100, 10), labels = function(x) paste0(x, "%")) +
  geom_text(size = 2.5,
             aes(y = perc + 3, label = paste0(round(perc, 1), "%"), group = party),
             position = position_dodge(width = 0.9)) +
  theme(legend.position = "bottom") +
  labs(x = "", y = "\nBudget Spend on Targeting Criterias (in % of Total Budget)", caption = "Source: Meta Ad Library and data compiled by Who Targets Me. Data Viz: Fabio Votta (@favstats).")#,
```


## Top Contested (Interest) Audiences  {.tabset .tabset-fade .tabset-pills}

Here, we show the *top most contested interest audiences*, i.e. where both Democrats and Republicans have spend considerable amounts of money.

### Nov 25th - 1st December 2022 (Last 7 Days)


```{r}
interest_targeting <- readRDS("runoff7/interest_targeting.rds")

```

```{r}
contested_dat <- interest_targeting %>% 
  filter(page_name %in% c("All Pro-Democrat Pages", "All Pro-Republican Pages")) %>% 
  filter(!is_exclusion) %>% 
  # filter(total_spend >= 40000) %>% 
  filter(total_spend >= 500) %>%
  add_count(value) %>% 
  filter(n == 2) %>% 
  group_by(value) %>% 
  mutate(total_spenderino = sum(total_spend)) %>% 
  mutate(perc = total_spend/total_spenderino) %>%
  mutate(value = str_remove_all(value, "INTERESTS: |DEMOGRAPHICS: |BEHAVIORS: ")) %>% 
  mutate(value = str_replace_all(value, " \\s*\\([^\\)]+\\)", ""))


the_order <- contested_dat %>% 
  filter(page_name == "All Pro-Republican Pages") %>% 
  arrange(desc(perc)) %>% 
  pull(value) %>% 
  unique()

lab_dat <- contested_dat %>% 
  # distinct(value, .keep_all = T) %>% 
  filter(page_name == "All Pro-Republican Pages") %>% 
  mutate(labb = paste0("$", scales::comma(round(total_spenderino)))) %>% 
  select(page_name, value, labb)


```

```{r, fig.height=10}
contested_dat %>% 
  # mutate(the_order = )
  left_join(lab_dat) %>% 
  mutate(value = factor(value, the_order)) %>% 
  mutate(page_name = factor(page_name, c("All Pro-Republican Pages", "All Pro-Democrat Pages"))) %>% 
  # filter()
  ggplot(aes(value, perc)) +
  geom_col(aes(fill = page_name), position = position_stack(), alpha = 0.8) +
  coord_flip() +
  geom_label(aes(label = labb),y=1.2,
            position = position_stack(vjust = 0.5),
            hjust = 1, label.size = NA,
            size = 4) + expand_limits(y = 1.2) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  scale_y_continuous(labels = scales::percent, breaks = c(0, 0.25, 0.5, 0.75, 1)) +
  scale_fill_manual("Page Groups", values = c("#ff2700", "#008fd5")) +
  ggthemes::theme_hc() +
  labs(x = "Targeting Criteria", y = "% of Budget Spend on Targeting Criteria", #title = "Most Contested Target Audiences in US 2022 Midterms", subtitle = " ", 
       caption = "Source: Meta Ad Library and data compiled by Who Targets Me. Data Viz: Fabio Votta (@favstats).") +
  theme(legend.position = "bottom", plot.title = element_text(size = 28, face = "bold", hjust = 1.25), text=element_text(family="mono"), plot.caption = element_text(hjust = -1.7))


```


### Nov 2nd - 1st December 2022 (Last 30 Days)

```{r}
interest_targeting30 <- readRDS("runoff/interest_targeting.rds")

```

```{r}
contested_dat <- interest_targeting30 %>% 
  filter(page_name %in% c("All Pro-Democrat Pages", "All Pro-Republican Pages")) %>% 
  filter(!is_exclusion) %>% 
  filter(total_spend >= 10000) %>%
  # filter(total_spend >= 500) %>%
  add_count(value) %>% 
  filter(n == 2) %>% 
  group_by(value) %>% 
  mutate(total_spenderino = sum(total_spend)) %>% 
  mutate(perc = total_spend/total_spenderino) %>%
  mutate(value = str_remove_all(value, "INTERESTS: |DEMOGRAPHICS: |BEHAVIORS: ")) %>% 
  mutate(value = str_replace_all(value, " \\s*\\([^\\)]+\\)", ""))


the_order <- contested_dat %>% 
  filter(page_name == "All Pro-Republican Pages") %>% 
  arrange(desc(perc)) %>% 
  pull(value) %>% 
  unique()

lab_dat <- contested_dat %>% 
  # distinct(value, .keep_all = T) %>% 
  filter(page_name == "All Pro-Republican Pages") %>% 
  mutate(labb = paste0("$", scales::comma(round(total_spenderino)))) %>% 
  select(page_name, value, labb)


```

```{r, fig.height=12}
contested_dat %>% 
  # mutate(the_order = )
  left_join(lab_dat) %>% 
  mutate(value = factor(value, the_order)) %>% 
  mutate(page_name = factor(page_name, c("All Pro-Republican Pages", "All Pro-Democrat Pages"))) %>% 
  # filter()
  ggplot(aes(value, perc)) +
  geom_col(aes(fill = page_name), position = position_stack(), alpha = 0.8) +
  coord_flip() +
  geom_label(aes(label = labb),y=1.25,
            position = position_stack(vjust = 0.5),
            hjust = 1, label.size = NA,
            size = 4) + expand_limits(y = 1.2) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  scale_y_continuous(labels = scales::percent, breaks = c(0, 0.25, 0.5, 0.75, 1)) +
  scale_fill_manual("Page Groups", values = c("#ff2700", "#008fd5")) +
  ggthemes::theme_hc() +
  labs(x = "Targeting Criteria", y = "% of Budget Spend on Targeting Criteria", #title = "Most Contested Target Audiences in US 2022 Midterms", subtitle = " ", 
       caption = "Source: Meta Ad Library and data compiled by Who Targets Me. Data Viz: Fabio Votta (@favstats).") +
  theme(legend.position = "bottom", plot.title = element_text(size = 28, face = "bold", hjust = 1.2), text=element_text(family="mono"), plot.caption = element_text(hjust = -1.7))


```

## Top Excluded Interests  {.tabset .tabset-fade .tabset-pills}

On Meta, advertisers can also **pay to exclude certain audiences**. The graph below shows which audiences are being **excluded** (i.e. how much money advertisers spend on ads that exclude audiences). Interestingly, Pro-Democrat pages seem to both include and exclude likely conservative leaning audiences (*NASCAR*, *Hunting*). Pro-Republican pages exclude people interested in *Veganism* and *Toyota Prius*. 

### Nov 25th - 1st December 2022 (Last 7 Days)

```{r, fig.width=10}
interest_targeting %>% 
  filter(is_exclusion) %>% 
  filter(is.na(page_id)) %>% 
  group_by(page_name) %>% 
  slice(1:20)  %>%
  mutate(value = str_remove_all(value, "INTERESTS: |DEMOGRAPHICS: |BEHAVIORS: ")) %>% 
  mutate(value = fct_reorder(value, total_spend)) %>% 
  ggplot(aes(value, total_spend, fill = page_name)) +
  geom_col() +
  facet_wrap(~page_name, scales = "free") +
  coord_flip()  +
  scale_fill_manual("Page Groups", values = c("#ff2700", "#008fd5")) +
  ggthemes::theme_hc() +
  theme(legend.position = "none") +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Interest Exlcusion", y = "\nTotal Money Spend on Excluding Audiences (in $)", 
       caption = "Source: Meta Ad Library and data compiled by Who Targets Me. Data Viz: Fabio Votta (@favstats).")
```


### Nov 2nd - 1st December 2022 (Last 30 Days)

```{r}

interest_targeting30 %>% 
  filter(is_exclusion) %>% 
  filter(is.na(page_id)) %>% 
  group_by(page_name) %>% 
  slice(1:20)  %>%
  mutate(value = str_remove_all(value, "INTERESTS: |DEMOGRAPHICS: |BEHAVIORS: ")) %>% 
  mutate(value = str_trunc(value, 20)) %>% 
  mutate(value = fct_reorder(value, total_spend)) %>% 
  ggplot(aes(value, total_spend, fill = page_name)) +
  geom_col() +
  facet_wrap(~page_name, scales = "free") +
  coord_flip()  +
  scale_fill_manual("Page Groups", values = c("#ff2700", "#008fd5")) +
  ggthemes::theme_hc() +
  theme(legend.position = "none") +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Interest Exlcusion", y = "\nTotal Money Spend on Excluding Audiences (in $)", 
       caption = "Source: Meta Ad Library and data compiled by Who Targets Me. Data Viz: Fabio Votta (@favstats).")
```


## ZIP Code Targeting  {.tabset .tabset-fade .tabset-pills}

Both Pro-Democrat and Pro-Republican pages **target voters in and around cities**, most notably Atlanta. 

### Nov 25th - 1st December 2022 (Last 7 Days)  {.tabset .tabset-fade .tabset-pills}

#### Pro-Democrat Pages


```{r}

```

```{r}
aggregate_zips <- readRDS("runoff7/aggregate_zips.rds")
zip_data <- readRDS("runoff7/zip_data.rds")
# shapename <- readRDS("data/statesdata/shapename.rds")
shapename <- sf::read_sf('geocoders/cb_2020_us_zcta520_500k.shp')
zipcounty <- readRDS("dashboard/zipcounty.rds")


library(scales)
library(leaflet)
library(shiny)
# debugonce(create_geo_dat)
create_geo_dat(aggregate_zips, 
               zip_data,
               "By Group",
               "Target", 
               "Georgia",
               "All Pro-Democrat Pages") %>% 
  leafletter()
```

#### Pro-Republican Pages

```{r}
create_geo_dat(aggregate_zips, 
               zip_data,
               "By Group",
               "Target", 
               "Georgia",
               "All Pro-Republican Pages") %>% 
  leafletter()
```


### Nov 2nd - 1st December 2022 (Last 30 Days) {.tabset .tabset-fade .tabset-pills}

#### Pro-Democrat Pages 


```{r}
aggregate_zips30 <- readRDS("runoff/aggregate_zips.rds")
zip_data30 <- readRDS("runoff/zip_data.rds")
# shapename <- readRDS("data/statesdata/shapename.rds")
shapename <- sf::read_sf('geocoders/cb_2020_us_zcta520_500k.shp')
zipcounty <- readRDS("dashboard/zipcounty.rds")


library(scales)
library(leaflet)
library(shiny)
# debugonce(create_geo_dat)
create_geo_dat(aggregate_zips30, 
               zip_data30,
               "By Group",
               "Target", 
               "Georgia",
               "All Pro-Democrat Pages") %>% 
  leafletter()
```



#### Pro-Republican Pages

```{r}
create_geo_dat(aggregate_zips30, 
               zip_data30,
               "By Group",
               "Target", 
               "Georgia",
               "All Pro-Republican Pages") %>% 
  leafletter()
```




## Gender Targeting {.tabset .tabset-fade .tabset-pills}

**Pro-Democrat pages spend much more money on targeting Women** compared to Pro-Republican pages, even compared to the overall lower budget of Pro-Republican pages.

### Nov 25th - 1st December 2022 (Last 7 Days)  

```{r}
gender_targeting <- readRDS("runoff7/gender_targeting.rds")
```
```{r}
gender_targeting %>%
      group_by(left_vs_right, value) %>% 
  summarise(total_spend = sum(total_spend)) %>% 
  drop_na() %>% 
  ggplot(aes(value, total_spend, fill = left_vs_right)) +
  geom_col(position = position_dodge())  +
  scale_fill_manual("Page Groups", values = c("#008fd5", "#ff2700")) +
  ggthemes::theme_hc() +
  labs(x = "Targeting Criteria", y = "Budget Spend on Gender Criteria", #title = "Most Contested Target Audiences in US 2022 Midterms", subtitle = " ", 
       caption = "Source: Meta Ad Library and data compiled by Who Targets Me. Data Viz: Fabio Votta (@favstats).") +
  geom_text(aes(y = total_spend + 1000, label = paste0("$", scales::comma(round(total_spend)))), position = position_dodge(width = 0.9))
```


### Nov 2nd - 1st December 2022 (Last 30 Days) 


```{r}
gender_targeting30 <- readRDS("runoff/gender_targeting.rds")
```
```{r}
gender_targeting30 %>%
      group_by(left_vs_right, value) %>% 
  summarise(total_spend = sum(total_spend)) %>% 
  drop_na() %>% 
  ggplot(aes(value, total_spend, fill = left_vs_right)) +
  geom_col(position = position_dodge())  +
  scale_fill_manual("Page Groups", values = c("#008fd5", "#ff2700")) +
  ggthemes::theme_hc() +
  labs(x = "", y = "Budget Spend on Gender Targeting", #title = "Most Contested Target Audiences in US 2022 Midterms", subtitle = " ", 
       caption = "Source: Meta Ad Library and data compiled by Who Targets Me. Data Viz: Fabio Votta (@favstats).") +
  geom_text(aes(y = total_spend + 50000, label = paste0("$", scales::comma(round(total_spend)))), position = position_dodge(width = 0.9))
```

## Language Targeting {.tabset .tabset-fade .tabset-pills}

Both Pro-Democrat and Pro-Republican pages run ads targeted at English-speakers. However, Pro-Democrat pages spend a lot of money on reaching out to Spanish speakers, much more than Republicans.


### Nov 25th - 1st December 2022 (Last 7 Days)  

```{r}
language_targeting <- readRDS("runoff7/language_targeting.rds")
```

```{r}
language_targeting %>%
      group_by(left_vs_right, value) %>% 
  summarise(total_spend = sum(total_spend)) %>% 
  drop_na() %>% 
    filter(total_spend >= 1000) %>% 
  ggplot(aes(value, total_spend, fill = left_vs_right)) +
  geom_col(position = position_dodge())  +
  scale_fill_manual("Page Groups", values = c("#008fd5", "#ff2700")) +
  ggthemes::theme_hc() +
  labs(x = "", y = "Budget Spend on Language Targeting", #title = "Most Contested Target Audiences in US 2022 Midterms", subtitle = " ", 
       caption = "Source: Meta Ad Library and data compiled by Who Targets Me. Data Viz: Fabio Votta (@favstats).") +
  geom_text(aes(y = total_spend + 1000, label = paste0("$", scales::comma(round(total_spend)))), position = position_dodge(width = 0.9))
```

### Nov 2nd - 1st December 2022 (Last 30 Days) 


```{r}
language_targeting30 <- readRDS("runoff/language_targeting.rds")
```

```{r}
language_targeting30 %>%
      group_by(left_vs_right, value) %>% 
  summarise(total_spend = sum(total_spend)) %>% 
  drop_na() %>% 
  filter(total_spend >= 10000) %>% 
  ggplot(aes(value, total_spend, fill = left_vs_right)) +
  geom_col(position = position_dodge())  +
  scale_fill_manual("Page Groups", values = c("#008fd5", "#ff2700")) +
  ggthemes::theme_hc() +
  labs(x = "", y = "Budget Spend on Language Targeting", #title = "Most Contested Target Audiences in US 2022 Midterms", subtitle = " ", 
       caption = "Source: Meta Ad Library and data compiled by Who Targets Me. Data Viz: Fabio Votta (@favstats).") +
  geom_text(aes(y = total_spend + 25000, label = paste0("$", scales::comma(round(total_spend)))), position = position_dodge(width = 0.9))
```


