---
title: "Georgia Run-Off 2022"
subtitle: "How Do Political Campaigns Target Voters?"
author: "Who Targets Me<br>Fabio Votta"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
    highlight: kate
---



```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(
    cache = T, 
    echo = F, 
    warning = F, 
    message = F, 
    cache.lazy = FALSE
)


# pacman::p_load(tidyverse, highcharter)
library(tidyverse)
library(highcharter)
library(gt)



options(scipen = 999)

source("helpers.R")

# pro_reps <- us_advertisers %>% 
#   filter(left_vs_right == "All Republican-supporting pages")

# source("utils.R")
```

```{r}
# here::here()
us_advertisers <- vroom::vroom("data/wtm-advertisers-us-2022-11-28T14_22_01.338Z.csv") %>%
  # count(entities.short_name, sort = T) %>%
  filter(entities.short_name %in% c("GOP", "Dems", "DemPACs", "RepPACs", "Con", "Prog")) %>%
  mutate(left_vs_right = ifelse(entities.short_name %in% c("GOP", "Con", "RepPACs"), "All Republican-supporting pages", "All Democrat-supporting pages")) %>%
  select(page_name = name, page_id = advertisers_platforms.advertiser_platform_ref, advertiser_type = entities.short_name, left_vs_right) %>%
  mutate(advertiser_type = case_when(
    advertiser_type == "Con" ~ "Conservatives (non-party, non-PACs)",
    advertiser_type == "Prog" ~ "Progressives  (non-party, non-PACs)",
    advertiser_type == "RepPACs" ~ "Republican PACs",
    advertiser_type == "DemPACs" ~ "Democrat PACs",
    advertiser_type == "GOP" ~ "GOP (candidates, parties)",
    advertiser_type == "Dems" ~ "Democrats (candidates, parties)"
  )) #%>% 

runoff <- readRDS("data/runoff7.rds")  %>%
  left_join(us_advertisers %>% mutate(internal_id = as.character(page_id))) %>%
  mutate(total_spend_formatted = readr::parse_number(total_spend_formatted)) %>%
  mutate(total_spend_formatted = ifelse(total_spend_formatted == 100, 1, total_spend_formatted)) %>%
  mutate(total_spend = total_spend_formatted * total_spend_pct)  %>% 
  mutate(page_name = case_when(
    str_detect(page_name, "All Pro-Republican Pages") ~ "All Republican-supporting pages",
        str_detect(page_name, "All Pro-Democrat Pages") ~ "All Democrat-supporting pages",
    T ~ page_name
  ))  

runoff30 <- readRDS("data/runoff.rds")  %>%
  left_join(us_advertisers %>% mutate(internal_id = as.character(page_id))) %>%
  mutate(total_spend_formatted = readr::parse_number(total_spend_formatted)) %>%
  mutate(total_spend_formatted = ifelse(total_spend_formatted == 100, 1, total_spend_formatted)) %>%
  mutate(total_spend = total_spend_formatted * total_spend_pct)    %>% 
  mutate(page_name = case_when(
    str_detect(page_name, "All Pro-Republican Pages") ~ "All Republican-supporting pages",
        str_detect(page_name, "All Pro-Democrat Pages") ~ "All Democrat-supporting pages",
    T ~ page_name
  ))  


# runoff <- runoff  %>%
#   left_join(us_advertisers %>% mutate(internal_id = as.character(page_id))) %>%
#   mutate(total_spend_formatted = readr::parse_number(total_spend_formatted)) %>%
#   mutate(total_spend_formatted = ifelse(total_spend_formatted == 100, 1, total_spend_formatted)) %>%
#   mutate(total_spend = total_spend_formatted * total_spend_pct)    %>% 
#   mutate(page_name = case_when(
#     str_detect(page_name, "All Pro-Republican Pages") ~ "All Republican-supporting pages",
#         str_detect(page_name, "All Pro-Democrat Pages") ~ "All Democrat-supporting pages",
#     T ~ page_name
#   ))  

# if(nrow(runoff30)!=119380) stop("WHAAAAAT") 

# runoff30 %>% count(ds, sort = T)
```


```{r}
# runoff30 %>%
#   distinct(internal_id, .keep_all = T) %>%
#   # mutate(spen = readr::parse_number(total_spend_formatted)) %>%
#   arrange(-total_spend_formatted) %>% 
#   select(internal_id, page_name, total_spend_formatted, left_vs_right, everything()) %>% 
#   filter(is.na(left_vs_right))
#   
# pro_demmm <- c("92021268117", "107483817554072", "19972147679", "6204742571")
# 
# pro_reppp <- c("127225910653607", "49651563727")
```


## Methodology

We coded over 3,000 US political advertisers during the 2022 midterm campaigns to better understand how campaigns use different targeting methods made available by Meta. To do this, we used data from the [Meta Ad Library](https://www.facebook.com/ads/library/), most often the new 'Audience' data which gives some detail on how pages target their ads.

To better understand the Georgia runoff election, we kept only advertisers who:

1. Advertised in the last 7 days (Nov 30th - Dec 6th 2022)
2. Advertised in the last 30 days (Nov 7th - Dec 6th 2022)


before election day.

> Note: Meta only provides 7, 30 and 90 days windows for the targeting data in their Ad Library. Meta's data also lags by a few days. We will update this report as soon as new data is available. 


## Topline statistics  {.tabset .tabset-fade .tabset-pills}

First, Democrat (Warnock)-supporting pages are clearly and consistently outspending Replubican (Walker)-supporting pages. Democrat-supporting pages have also run many more ads.


### Nov 30th - Dec 6th 2022 (Last 7 Days)

```{r}
# runoff %>% count(ds)
```


```{r}
runoff %>% 
  distinct(internal_id, .keep_all = T) %>% 
  group_by(left_vs_right) %>% 
  summarize(total_num_ads = n()) %>% 
  drop_na() %>% 
    mutate(total_num_ads = scales::comma(total_num_ads)) %>%
  pivot_wider(names_from = left_vs_right, values_from = total_num_ads) %>% 
  mutate(type = "Total Number of Advertisers") %>% 
  
  
  bind_rows(
    runoff %>% 
  distinct(internal_id, .keep_all = T) %>% 
  group_by(left_vs_right) %>% 
  arrange(desc(total_spend_formatted)) %>% 
    slice(1:3) %>% 
    mutate(total_spend_formatted = scales::comma(total_spend_formatted)) %>%
    mutate(lab = paste0(page_name, " ($", total_spend_formatted, ")")) %>% 
    select(left_vs_right, lab) %>% 
  drop_na() %>% 
  summarize(lab = paste0("<br>", 1:3, ". ", lab, collapse = "")) %>% 
  pivot_wider(names_from = left_vs_right, values_from = lab) %>% 
  mutate(type = "Top Advertiser")
  ) %>% 
  
  bind_rows(
    runoff %>% 
  distinct(internal_id, .keep_all = T) %>% 
  group_by(left_vs_right) %>% 
  summarize(total_num_ads = sum(total_num_ads)) %>% 
  drop_na() %>% 
    mutate(total_num_ads = scales::comma(total_num_ads)) %>% 
  pivot_wider(names_from = left_vs_right, values_from = total_num_ads) %>% 
  mutate(type = "Total Number of Ads")
  )%>% 
  bind_rows(
    runoff %>% 
  distinct(internal_id, .keep_all = T) %>% 
  group_by(left_vs_right) %>% 
  summarize(total_spend_formatted = sum(total_spend_formatted)) %>% 
    mutate(total_spend_formatted = scales::comma(total_spend_formatted)) %>% 
  drop_na() %>% 
  pivot_wider(names_from = left_vs_right, values_from = total_spend_formatted) %>% 
  mutate(type = "Total Spend ($)") 
  ) %>% 



  gt(
    rowname_col = "type"
    # groupname_col = "group"
  ) %>% 
  fmt_markdown(columns = everything())
```

### Nov 7th - Dec 6th 2022 (Last 30 Days)

```{r}
runoff30 %>% 
  distinct(internal_id, .keep_all = T) %>% 
  group_by(left_vs_right) %>% 
  summarize(total_num_ads = n()) %>% 
  drop_na() %>% 
    mutate(total_num_ads = scales::comma(total_num_ads)) %>%
  pivot_wider(names_from = left_vs_right, values_from = total_num_ads) %>% 
  mutate(type = "Total Number of Advertisers") %>% 
  
  
  bind_rows(
    runoff30 %>% 
  distinct(internal_id, .keep_all = T) %>% 
  group_by(left_vs_right) %>% 
  arrange(desc(total_spend_formatted)) %>% 
    slice(1:3) %>% 
    mutate(total_spend_formatted = scales::comma(total_spend_formatted)) %>%
    mutate(lab = paste0(page_name, " ($", total_spend_formatted, ")")) %>% 
    select(left_vs_right, lab) %>% 
  drop_na() %>% 
  summarize(lab = paste0("<br>", 1:3, ". ", lab, collapse = "")) %>% 
  pivot_wider(names_from = left_vs_right, values_from = lab) %>% 
  mutate(type = "Top Advertiser")
  ) %>% 
  
  bind_rows(
    runoff30 %>% 
  distinct(internal_id, .keep_all = T) %>% 
  group_by(left_vs_right) %>% 
  summarize(total_num_ads = sum(total_num_ads)) %>% 
  drop_na() %>% 
    mutate(total_num_ads = scales::comma(total_num_ads)) %>% 
  pivot_wider(names_from = left_vs_right, values_from = total_num_ads) %>% 
  mutate(type = "Total Number of Ads")
  )%>% 
  bind_rows(
    runoff30 %>% 
  distinct(internal_id, .keep_all = T) %>% 
  group_by(left_vs_right) %>% 
  summarize(total_spend_formatted = sum(total_spend_formatted)) %>% 
    mutate(total_spend_formatted = scales::comma(total_spend_formatted)) %>% 
  drop_na() %>% 
  pivot_wider(names_from = left_vs_right, values_from = total_spend_formatted) %>% 
  mutate(type = "Total Spend") 
  ) %>% 



  gt(
    rowname_col = "type"
    # groupname_col = "group"
  ) %>% 
  fmt_markdown(columns = everything())
```



## Spending, broken down by targeting method  {.tabset .tabset-fade .tabset-pills}


How much did campaigns spend on different targeting methods? 

**Note that most advertisers spend the vast majority of their budget (~65%+) on Meta's Custom Audiences.**

### Nov 30th - Dec 6th 2022 (Last 7 Days)


```{r}

pro_reps <- us_advertisers %>% 
  filter(left_vs_right == "All Republican-supporting pages")

pro_dems <- us_advertisers %>% 
  filter(left_vs_right == "All Democrat-supporting pages")


runoff_spend_reps <- calculate_spend_on_each(runoff %>% filter(internal_id %in% pro_reps$page_id))

runoff_spend_dems <- calculate_spend_on_each(runoff %>% filter(internal_id %in% pro_dems$page_id))

runoff_spend_reps %>% 
  mutate(party= "Pro-Republican Pages") %>% 
  bind_rows(runoff_spend_dems %>% 
  mutate(party= "Pro-Democrat Pages")) %>% 
  filter(perc >= 0.1) %>%
  add_count(target) %>% 
  filter(n == 2) %>% 
  mutate(target = case_when(
    target == "custom_audience" ~ "Custom Audiences",
    target == "countries" ~ "GEOGRAPHY: Entire US",
    target == "regions" ~ "GEOGRAPHY: State(s)",
    target == "lookalike_audience" ~ "Lookalike Audiences",
    target == "interest" ~ "Interests",
    target == "age" ~ "Age",
    target == "zips" ~ "GEOGRAPHY: ZIP-Codes",
    target == "CITY" ~ "GEOGRAPHY: Cities",
    target == "language" ~ "Language",
    target == "gender" ~ "Gender",
    target == "electoral_districts" ~ "GEOGRAPHY: Electoral Districts",
    target == "COUNTY" ~ "GEOGRAPHY: Counties",
  )) %>% 
  mutate(target = fct_reorder(target, perc)) %>% 
  ggplot(aes(target, perc)) +
  geom_col(position = position_dodge(width = 0.9), aes(fill = party)) +
  coord_flip() +
  scale_fill_manual(name = "Pages", values = c("#008fd5", "#ff2700")) +
  theme_minimal() +
  scale_y_continuous(breaks = seq(0, 100, 10), labels = function(x) paste0(x, "%")) +
  geom_text(size = 2.5,
             aes(y = perc + 3, label = paste0(round(perc, 1), "%"), group = party),
             position = position_dodge(width = 0.9)) +
  theme(legend.position = "bottom") +
  labs(x = "", y = "\nBudget spent on targeting method (% of Total spend)", caption = "Source: Meta Ad Library using data compiled by Who Targets Me. Data Viz: Fabio Votta (@favstats).")#,
       # title = "How much did campaigns spend on particular targeting methods?",
       # subtitle = "Period: Nov 30th - Dec 6th 2022 (Last 7 Days)")

```

### Nov 7th - Dec 6th 2022 (Last 30 Days)

```{r}
runoff30_spend_reps <- calculate_spend_on_each(runoff30 %>% filter(internal_id %in% pro_reps$page_id))

runoff30_spend_dems <- calculate_spend_on_each(runoff30 %>% filter(internal_id %in% pro_dems$page_id))

runoff30_spend_reps %>% 
  mutate(party= "Republican-supporting pages") %>% 
  bind_rows(runoff30_spend_dems %>% 
  mutate(party= "Democrat-supporting pages")) %>% 
  filter(perc >= 0.1) %>%
  add_count(target) %>% 
  filter(n == 2) %>% 
  mutate(target = case_when(
    target == "custom_audience" ~ "Custom Audiences",
    target == "countries" ~ "GEOGRAPHY: Entire US",
    target == "regions" ~ "GEOGRAPHY: State(s)",
    target == "lookalike_audience" ~ "Lookalike Audiences",
    target == "interest" ~ "Interests",
    target == "age" ~ "Age",
    target == "zips" ~ "GEOGRAPHY: ZIP-Codes",
    target == "CITY" ~ "GEOGRAPHY: Cities",
    target == "language" ~ "Language",
    target == "gender" ~ "Gender",
    target == "electoral_districts" ~ "GEOGRAPHY: Electoral Districts",
    target == "COUNTY" ~ "GEOGRAPHY: Counties",
  )) %>% 
  mutate(target = fct_reorder(target, perc)) %>% 
  drop_na() %>% 
  ggplot(aes(target, perc)) +
  geom_col(position = position_dodge(width = 0.9), aes(fill = party)) +
  coord_flip() +
  scale_fill_manual(name = "Pages", values = c("#008fd5", "#ff2700")) +
  theme_minimal() +
  scale_y_continuous(breaks = seq(0, 100, 10), labels = function(x) paste0(x, "%")) +
  geom_text(size = 2.5,
             aes(y = perc + 3, label = paste0(round(perc, 1), "%"), group = party),
             position = position_dodge(width = 0.9)) +
  theme(legend.position = "bottom") +
  labs(x = "", y = "\nBudget Spend on Targeting Criterias (in % of Total Budget)", caption = "Source: Meta Ad Library and data compiled by Who Targets Me. Data Viz: Fabio Votta (@favstats).")#,
```


## Top Contested (Interest) Audiences  {.tabset .tabset-fade .tabset-pills}

Here, we show the *top most contested interest audiences*, i.e. where both Democrats and Republicans have spent considerable amounts of money competing to reach voters with the same interests.

### Nov 30th - Dec 6th 2022 (Last 7 Days)


```{r}
interest_targeting <- readRDS("runoff7/interest_targeting.rds")   %>% 
  mutate(page_name = case_when(
    str_detect(page_name, "All Pro-Republican Pages") ~ "All Republican-supporting pages",
        str_detect(page_name, "All Pro-Democrat Pages") ~ "All Democrat-supporting pages",
    T ~ page_name
  ))  

```

```{r}
contested_dat <- interest_targeting %>% 
  filter(page_name %in% c("All Democrat-supporting pages", "All Republican-supporting pages")) %>% 
  filter(!is_exclusion) %>% 
  # filter(total_spend >= 40000) %>% 
  filter(total_spend >= 500) %>%
  add_count(value) %>% 
  filter(n == 2) %>% 
  group_by(value) %>% 
  mutate(total_spenderino = sum(total_spend)) %>% 
  mutate(perc = total_spend/total_spenderino) %>%
  mutate(value = str_remove_all(value, "INTERESTS: |DEMOGRAPHICS: |BEHAVIORS: ")) %>% 
  mutate(value = str_replace_all(value, " \\s*\\([^\\)]+\\)", ""))


the_order <- contested_dat %>% 
  filter(page_name == "All Republican-supporting pages") %>%   arrange(desc(perc)) %>% 
  pull(value) %>% 
  unique()

lab_dat <- contested_dat %>% 
  # distinct(value, .keep_all = T) %>% 
  filter(page_name == "All Republican-supporting pages") %>% 
  mutate(labb = paste0("$", scales::comma(round(total_spenderino)))) %>% 
  select(page_name, value, labb)


```

```{r, fig.width  = 10, fig.height=10}
contested_dat %>% 
  # mutate(the_order = )
  left_join(lab_dat) %>% 
  mutate(value = factor(value, the_order)) %>% 
  mutate(page_name = factor(page_name, c("All Republican-supporting pages", "All Democrat-supporting pages"))) %>% 
  # filter()
  ggplot(aes(value, perc)) +
  geom_col(aes(fill = page_name), position = position_stack(), alpha = 0.8) +
  coord_flip() +
  geom_label(aes(label = labb),y=1.225,
            position = position_stack(vjust = 0.5),
            hjust = 1, label.size = NA,
            size = 4) + expand_limits(y = 1.2) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  scale_y_continuous(labels = scales::percent, breaks = c(0, 0.25, 0.5, 0.75, 1)) +
  scale_fill_manual("Page Groups", values = c("#ff2700", "#008fd5")) +
  ggthemes::theme_hc() +
  labs(x = "Targeting criteria", y = "% of budget spent on targeting method", #title = "Contested targeted interests (US midterms 2022)", subtitle = " ", 
       caption = "Source: Meta Ad Library and data compiled by Who Targets Me. Data Viz: Fabio Votta (@favstats).") +
  theme(legend.position = "bottom", plot.title = element_text(size = 28, face = "bold", hjust = 1.25), text=element_text(family="mono"), plot.caption = element_text(hjust = -1.7)) +
  guides(fill=guide_legend(nrow=2,byrow=TRUE))


```


### Nov 7th - Dec 6th 2022 (Last 30 Days)

```{r}
interest_targeting30 <- readRDS("runoff/interest_targeting.rds")   %>% 
  mutate(page_name = case_when(
    str_detect(page_name, "All Pro-Republican Pages") ~ "All Republican-supporting pages",
        str_detect(page_name, "All Pro-Democrat Pages") ~ "All Democrat-supporting pages",
    T ~ page_name
  ))

# interest_targeting30 %>% count(page_name, sort = T)
```

```{r}
contested_dat <- interest_targeting30 %>% 
  filter(page_name %in% c("All Democrat-supporting pages", "All Republican-supporting pages")) %>% 
  filter(!is_exclusion) %>% 
  filter(total_spend >= 10000) %>%
  # filter(total_spend >= 500) %>%
  add_count(value) %>% 
  filter(n == 2) %>% 
  group_by(value) %>% 
  mutate(total_spenderino = sum(total_spend)) %>% 
  mutate(perc = total_spend/total_spenderino) %>%
  mutate(value = str_remove_all(value, "INTERESTS: |DEMOGRAPHICS: |BEHAVIORS: ")) %>% 
  mutate(value = str_replace_all(value, " \\s*\\([^\\)]+\\)", ""))


the_order <- contested_dat %>% 
  filter(page_name == "All Republican-supporting pages") %>%   arrange(desc(perc)) %>% 
  pull(value) %>% 
  unique()

lab_dat <- contested_dat %>% 
  # distinct(value, .keep_all = T) %>% 
  filter(page_name == "All Republican-supporting pages") %>%   mutate(labb = paste0("$", scales::comma(round(total_spenderino)))) %>% 
  select(page_name, value, labb)


```

```{r, fig.width  = 8, fig.height=8}
contested_dat %>% 
  # mutate(the_order = )
  left_join(lab_dat) %>% 
  mutate(value = factor(value, the_order))  %>% 
  mutate(page_name = factor(page_name, c("All Republican-supporting pages", "All Democrat-supporting pages"))) %>% 
  # count(page_name, sort  =T)
  # filter()
  ggplot(aes(value, perc)) +
  geom_col(aes(fill = page_name), position = position_stack(), alpha = 0.8) +
  coord_flip() +
  geom_label(aes(label = labb),y=1.25,
            position = position_stack(vjust = 0.5),
            hjust = 1, label.size = NA,
            size = 4) + expand_limits(y = 1.2) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  scale_y_continuous(labels = scales::percent, breaks = c(0, 0.25, 0.5, 0.75, 1)) +
  scale_fill_manual("Page Groups", values = c("#ff2700", "#008fd5")) +
  ggthemes::theme_hc() +
  labs(x = "Targeting Criteria", y = "% of budget spent on targeting method", #title = "Contested targeted interests (US midterms 2022)", subtitle = " ", 
       caption = "Source: Meta Ad Library and data compiled by Who Targets Me. Data Viz: Fabio Votta (@favstats).") +
  theme(legend.position = "bottom", plot.title = element_text(size = 28, face = "bold", hjust = 1.2), text=element_text(family="mono"), plot.caption = element_text(hjust = -1.7))  +
  guides(fill=guide_legend(nrow=2,byrow=TRUE))


```

## Top Excluded Interests  {.tabset .tabset-fade .tabset-pills}

On Meta, advertisers can also **choose to exclude certain audiences from seeing their ads**. The graph below shows which audiences are being **excluded** and how much pages are paying for ads that exclude them. Interestingly, Democrat-supporting pages seem to both include and exclude likely conservative leaning audiences (*NASCAR*, *Hunting*). Pro-Republican pages exclude people interested in *Veganism* and *Toyota Prius*. 

### Nov 30th - Dec 6th 2022 (Last 7 Days)

```{r, fig.height=6, fig.width=10}
interest_targeting %>% 
  filter(is_exclusion) %>% 
  filter(is.na(page_id)) %>% 
  mutate(value = str_remove_all(value, "INTERESTS: |DEMOGRAPHICS: |BEHAVIORS: ")) %>% 
  mutate(value = str_trunc(value, 30)) %>% 
  group_by(page_name)  %>% 
  arrange(desc(total_spend)) %>% 
  distinct(value, .keep_all = T) %>% 
  slice(1:20)  %>%
  ungroup() %>% 
  mutate(value = fct_reorder(value, total_spend)) %>% 
  ggplot(aes(value, total_spend, fill = page_name)) +
  geom_col() +
  facet_wrap(~page_name, scales = "free") +
  coord_flip()  +
  scale_fill_manual("Page Groups", values = c("#008fd5", "#ff2700")) +
  ggthemes::theme_hc() +
  theme(legend.position = "none") +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Interest Exlcusion", y = "\nTotal spent on ads that exclude audience (in $)", 
       caption = "Source: Meta Ad Library and data compiled by Who Targets Me. Data Viz: Fabio Votta (@favstats).") +
	scale_y_continuous(labels = scales::label_number(suffix = "k", scale = 1e-3)) 
```


### Nov 7th - Dec 6th 2022 (Last 30 Days)

```{r, fig.height=6, fig.width=10}
# print(interest_targeting30 %>% 
#   filter(is_exclusion) %>% 
#   filter(is.na(page_id)))

interest_targeting30 %>% 
  filter(is_exclusion) %>% 
  filter(is.na(page_id)) %>% 
  mutate(value = str_remove_all(value, "INTERESTS: |DEMOGRAPHICS: |BEHAVIORS: ")) %>%
  mutate(value = str_trunc(value, 30)) %>% 
  group_by(page_name)  %>% 
  arrange(desc(total_spend)) %>% 
  distinct(value, .keep_all = T) %>% 
  slice(1:20)  %>%
  ungroup() %>% 
  mutate(value = fct_reorder(value, total_spend)) %>% 
  ggplot(aes(value, total_spend, fill = page_name)) +
  geom_col() +
  facet_wrap(~page_name, scales = "free") +
  coord_flip()  +
  scale_fill_manual("Page Groups", values = c("#008fd5", "#ff2700")) +
  ggthemes::theme_hc() +
  theme(legend.position = "none") +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Interest Exlcusion", y = "\nTotal spent excluding audiences (in $)",
       caption = "Source: Meta Ad Library and data compiled by Who Targets Me. Data Viz: Fabio Votta (@favstats).")  +
	scale_y_continuous(labels = scales::label_number(suffix = "k", scale = 1e-3)) 
```


## ZIP Code Targeting  {.tabset .tabset-fade .tabset-pills}

Both Democrat-supporting and Republican-supporting pages **target voters in and around cities**, most notably Atlanta. 

### Nov 30th - Dec 6th 2022 (Last 7 Days)  {.tabset .tabset-fade .tabset-pills}

#### Democrat-supporting pages


```{r}

```

```{r}
aggregate_zips <- readRDS("runoff7/aggregate_zips.rds") 
zip_data <- readRDS("runoff7/zip_data.rds")
# shapename <- readRDS("data/statesdata/shapename.rds")
shapename <- sf::read_sf('geocoders/cb_2020_us_zcta520_500k.shp')
zipcounty <- readRDS("dashboard/zipcounty.rds")


library(scales)
library(leaflet)
library(shiny)
# debugonce(create_geo_dat)
create_geo_dat(aggregate_zips, 
               zip_data,
               "By Group",
               "Target", 
               "Georgia",
               "All Pro-Democrat Pages") %>% 
  leafletter()
```

#### Republican-supporting pages

```{r}
create_geo_dat(aggregate_zips,
               zip_data,
               "By Group",
               "Target",
               "Georgia",
               "All Pro-Republican Pages") %>%
  leafletter()
```


### Nov 7th - Dec 6th 2022 (Last 30 Days) {.tabset .tabset-fade .tabset-pills}

#### Democrat-supporting pages


```{r}
aggregate_zips30 <- readRDS("runoff/aggregate_zips.rds")
zip_data30 <- readRDS("runoff/zip_data.rds")
# shapename <- readRDS("data/statesdata/shapename.rds")
shapename <- sf::read_sf('geocoders/cb_2020_us_zcta520_500k.shp')
zipcounty <- readRDS("dashboard/zipcounty.rds")


library(scales)
library(leaflet)
library(shiny)
# debugonce(create_geo_dat)
create_geo_dat(aggregate_zips30, 
               zip_data30,
               "By Group",
               "Target", 
               "Georgia",
               "All Pro-Democrat Pages") %>% 
  leafletter()
```



#### Republican-supporting pages

```{r}
create_geo_dat(aggregate_zips30, 
               zip_data30,
               "By Group",
               "Target", 
               "Georgia",
               "All Pro-Republican Pages") %>% 
  leafletter()
```




## How the parties use gender to target ads {.tabset .tabset-fade .tabset-pills}

In the final days of the election campaign, **Republican-supporting pages spend more money than Democrat-supporting pages**. However, this trend reverses when looking at the 30-days window before the election: democrat-supporting pages spend more money targeting ads at women versus Republican-supporting pages, even factoring in their lower budget.

### Nov 30th - Dec 6th 2022 (Last 7 Days)  

```{r}
gender_targeting <- readRDS("runoff7/gender_targeting.rds")   %>% 
  mutate(page_name = case_when(
    str_detect(page_name, "All Pro-Republican Pages") ~ "All Republican-supporting pages",
        str_detect(page_name, "All Pro-Democrat Pages") ~ "All Democrat-supporting pages",
    T ~ page_name
  ))    %>% 
  mutate(left_vs_right = case_when(
    str_detect(left_vs_right, "All Pro-Republican Pages") ~ "All Republican-supporting pages",
        str_detect(left_vs_right, "All Pro-Democrat Pages") ~ "All Democrat-supporting pages",
    T ~ left_vs_right
  ))

# gender_targeting %>%
#   arrange(desc(total_spend))
```
```{r}
gender_targeting %>%
      group_by(left_vs_right, value) %>% 
  summarise(total_spend = sum(total_spend)) %>% 
  drop_na() %>% 
  ggplot(aes(value, total_spend, fill = left_vs_right)) +
  geom_col(position = position_dodge())  +
  scale_fill_manual("Page Groups", values = c("#008fd5", "#ff2700")) +
  ggthemes::theme_hc() +
  labs(x = "Targeting criteria", y = "Amount spent on targeting by gender", #title = "Most Contested Target Audiences in US 2022 Midterms", subtitle = " ", 
       caption = "Source: Meta Ad Library and data compiled by Who Targets Me. Data Viz: Fabio Votta (@favstats).") +
  geom_text(aes(y = total_spend + 1000, label = paste0("$", scales::comma(round(total_spend)))), position = position_dodge(width = 0.9)) +
  scale_y_continuous(labels= scales::comma)
```


### Nov 7th - Dec 6th 2022 (Last 30 Days) 


```{r}
gender_targeting30 <- readRDS("runoff/gender_targeting.rds") %>% 
  mutate(page_name = case_when(
    str_detect(page_name, "All Pro-Republican Pages") ~ "All Republican-supporting pages",
        str_detect(page_name, "All Pro-Democrat Pages") ~ "All Democrat-supporting pages",
    T ~ page_name
  ))    %>% 
  mutate(left_vs_right = case_when(
    str_detect(left_vs_right, "All Pro-Republican Pages") ~ "All Republican-supporting pages",
        str_detect(left_vs_right, "All Pro-Democrat Pages") ~ "All Democrat-supporting pages",
    T ~ left_vs_right
  ))
```
```{r}
gender_targeting30 %>%
      group_by(left_vs_right, value) %>% 
  summarise(total_spend = sum(total_spend)) %>% 
  drop_na() %>% 
  ggplot(aes(value, total_spend, fill = left_vs_right)) +
  geom_col(position = position_dodge())  +
  scale_fill_manual("Page Groups", values = c("#008fd5", "#ff2700")) +
  ggthemes::theme_hc() +
  labs(x = "", y = "Amount spent on targeting by gender", #title = "Most Contested Target Audiences in US 2022 Midterms", subtitle = " ", 
       caption = "Source: Meta Ad Library and data compiled by Who Targets Me. Data Viz: Fabio Votta (@favstats).") +
  geom_text(aes(y = total_spend + 20000, label = paste0("$", scales::comma(round(total_spend)))), position = position_dodge(width = 0.9))  +
  scale_y_continuous(labels= scales::comma)
```

## Language Targeting {.tabset .tabset-fade .tabset-pills}

Both Democrat and Republican ads spent most money on ads targeted at English-speakers. Democratic pages targeted spanish-speaking voters at a much higher rate.


### Nov 30th - Dec 6th 2022 (Last 7 Days)  

```{r}
language_targeting <- readRDS("runoff7/language_targeting.rds") %>% 
  mutate(page_name = case_when(
    str_detect(page_name, "All Pro-Republican Pages") ~ "All Republican-supporting pages",
        str_detect(page_name, "All Pro-Democrat Pages") ~ "All Democrat-supporting pages",
    T ~ page_name
  ))    %>% 
  mutate(left_vs_right = case_when(
    str_detect(left_vs_right, "All Pro-Republican Pages") ~ "All Republican-supporting pages",
        str_detect(left_vs_right, "All Pro-Democrat Pages") ~ "All Democrat-supporting pages",
    T ~ left_vs_right
  ))
```

```{r}
language_targeting %>%
      group_by(left_vs_right, value) %>% 
  summarise(total_spend = sum(total_spend)) %>% 
  drop_na() %>% 
    filter(total_spend >= 1000) %>% 
  ggplot(aes(value, total_spend, fill = left_vs_right)) +
  geom_col(position = position_dodge())  +
  scale_fill_manual("Page Groups", values = c("#008fd5", "#ff2700")) +
  ggthemes::theme_hc() +
  labs(x = "", y = "Total spent targeting by language", #title = "Most Contested Target Audiences in US 2022 Midterms", subtitle = " ", 
       caption = "Source: Meta Ad Library and data compiled by Who Targets Me. Data Viz: Fabio Votta (@favstats).") +
  geom_text(aes(y = total_spend + 2000, label = paste0("$", scales::comma(round(total_spend)))), position = position_dodge(width = 0.9))  +
  scale_y_continuous(labels= scales::comma)
```

### Nov 7th - Dec 6th 2022 (Last 30 Days) 


```{r}
language_targeting30 <- readRDS("runoff/language_targeting.rds")  %>% 
  mutate(page_name = case_when(
    str_detect(page_name, "All Pro-Republican Pages") ~ "All Republican-supporting pages",
        str_detect(page_name, "All Pro-Democrat Pages") ~ "All Democrat-supporting pages",
    T ~ page_name
  ))  %>% 
  mutate(left_vs_right = case_when(
    str_detect(left_vs_right, "All Pro-Republican Pages") ~ "All Republican-supporting pages",
        str_detect(left_vs_right, "All Pro-Democrat Pages") ~ "All Democrat-supporting pages",
    T ~ left_vs_right
  ))
```

```{r}
language_targeting30 %>%
      group_by(left_vs_right, value) %>% 
  summarise(total_spend = sum(total_spend)) %>% 
  drop_na() %>% 
  filter(total_spend >= 10000) %>% 
  ggplot(aes(value, total_spend, fill = left_vs_right)) +
  geom_col(position = position_dodge())  +
  scale_fill_manual("Page Groups", values = c("#008fd5", "#ff2700")) +
  ggthemes::theme_hc() +
  labs(x = "", y = "Total spent targeting by language", #title = "Most Contested Target Audiences in US 2022 Midterms", subtitle = " ", 
       caption = "Source: Meta Ad Library and data compiled by Who Targets Me. Data Viz: Fabio Votta (@favstats).") +
  geom_text(aes(y = total_spend + 15000, label = paste0("$", scales::comma(round(total_spend)))), position = position_dodge(width = 0.9))  +
  scale_y_continuous(labels= scales::comma)
```


